---
title: "Resit thesis"
author: "Juan Diego Tabares Guzman"
date: "2023-08-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Clear workspace

```{r}
rm(list = ls())
```

Prepare environment

```{r}
setwd("C:/Users/juant/OneDrive/Documentos/JD/Personal/Thesis/Drafts/Resit/Data")
```

Loading required packages

```{r, warning=FALSE}
library(readxl)
library(ggplot2)
library(gplots)
library(vars)
library(wpa)
library(ggpubr)
#remotes::install_github("angusmoore/varexternalinstrument", force = TRUE)
library(varexternalinstrument)
library(sos)
library(simts)
library(broom)

library(devtools)
#install_github("martinbaumgaertner/varexternal")
library(varexternal)
library(aTSA)
library(stargazer)
library(lpirfs)
```

Importing dataset for the estimation section 

```{r}
data <- read_excel("C:/Users/juant/OneDrive/Documentos/JD/Personal/Thesis/Drafts/Resit/Data/data_resit.xlsx")

data$us_cpi_inf <- data$us_cpi/100

data$real_oil_price <- data$refiner_acq_cost_comp/data$us_cpi_inf

data_var <- data.frame(log(data$oil_prod), log(data$world_ind_prod), log(data$real_oil_price), log(data$oecd_oil_stock), log(data$ind_prod_eu), log(data$ea19_cpi))

data_var <- data_var[-1:-18,]

colnames(data_var) <- c("oil_prod", "world_ind_prod", "real_oil_price", "oecd_oil_stock", "ind_prod_ea", "cpi_ea")

date = seq(from = as.Date("1975-07-01"), to = as.Date("2022-12-01"), by = 'month')

data_var$date <- date


data_var_t <- ts(data_var,start=c(1975,7),frequency=12)

head(data_var_t)
```
Graphs in levels

```{r}

oil_prod_g <- ggplot(data_var, aes(x=date, y=oil_prod)) +
  geom_line(color = "navy", size = 1) +
  theme_bw() +
  labs(x="Date", y="World oil prod. (MMBD)") +
  theme(text=element_text(size=12,  family="serif"))

world_ind_prod_g <- ggplot(data_var, aes(x=date, y=world_ind_prod)) +
  geom_line(color = "navy", size = 1) +
  theme_bw() +
  labs(x="Date", y="World ind. prod. (index)") +
  theme(text=element_text(size=12,  family="serif"))

real_oil_price_g <- ggplot(data_var, aes(x=date, y=real_oil_price)) +
  geom_line(color = "navy", size = 1) +
  theme_bw() +
  labs(x="Date", y="Real oil price (USD)") +
  theme(text=element_text(size=12,  family="serif"))

global_oil_stock_g <- ggplot(data_var, aes(x=date, y=oecd_oil_stock)) +
  geom_line(color = "navy", size = 1) +
  theme_bw() +
  labs(x="Date", y="World oil stocks (MMB)") +
  theme(text=element_text(size=12,  family="serif"))

oil_market_g1 <- ggarrange(oil_prod_g, world_ind_prod_g, ncol = 1, nrow = 2)
oil_market_g2 <- ggarrange(real_oil_price_g, global_oil_stock_g, ncol = 1, nrow = 2)

oil_market_g1
oil_market_g2

ind_prod_ea_g <- ggplot(data_var, aes(x=date, y=ind_prod_ea)) +
  geom_line(color = "navy", size = 1) +
  theme_bw() +
  labs(x="Date", y="EA Ind. prod. (index)") +
  theme(text=element_text(size=12,  family="serif"))

cpi_ea_g <- ggplot(data_var, aes(x=date, y=cpi_ea)) +
  geom_line(color = "navy", size = 1) +
  theme_bw() +
  labs(x="Date", y="EA CPI (index)") +
  theme(text=element_text(size=12,  family="serif"))

macro_g <- ggarrange(ind_prod_ea_g, cpi_ea_g, ncol = 1, nrow = 2)
macro_g

data_var_surprises <- data.frame(data$oil_supply_surprise, data$Date)
colnames(data_var_surprises) <- c("oil_supply_surprise", "date")
data_var_surprises <- data_var_surprises[-1:-111,]

surprise_g <- ggplot(data_var_surprises, aes(x=date, y=oil_supply_surprise)) +
  geom_line(color = "navy", size = 1) +
  theme_bw() +
  labs(x="Date", y="Revision in oil price expectations") +
  theme(text=element_text(size=12,  family="serif"))

```

Unit root test 

```{r}

## DF test oil production 

arima_oil_prod <- arima(data_var$oil_prod, order = c(8,0,0))
LjungBox(arima_oil_prod$residuals, lag=seq(1,25,1))

DF_oil_prod = ur.df(data_var$oil_prod, type = c("trend"), lags = 7)
summary(DF_oil_prod)

DF_oil_prod_inter = ur.df(data_var$oil_prod, type = c("drift"), lags = 7)
summary(DF_oil_prod_inter)

DF_oil_prod_none = ur.df(data_var$oil_prod, type = c("none"), lags = 7)
summary(DF_oil_prod_none)

## DF test world ind. prod.

arima_world_ind_prod <- arima(data_var$world_ind_prod, order = c(4,0,0))
LjungBox(arima_world_ind_prod$residuals, lag=seq(1,25,1))

DF_world_ind_prod = ur.df(data_var$world_ind_prod, type = c("trend"), lags = 3)
summary(DF_world_ind_prod)

DF_world_ind_prod_inter = ur.df(data_var$world_ind_prod, type = c("drift"), lags = 3)
summary(DF_world_ind_prod_inter)

DF_world_ind_prod_none = ur.df(data_var$world_ind_prod, type = c("none"), lags = 3)
summary(DF_world_ind_prod_none)


## DF test real oil price

arima_real_oil_price <- arima(data_var$real_oil_price, order = c(3,0,0))
LjungBox(arima_real_oil_price$residuals, lag=seq(1,25,1))

DF_real_oil_price = ur.df(data_var$real_oil_price, type = c("trend"), lags = 2)
summary(DF_real_oil_price)

DF_real_oil_price_inter = ur.df(data_var$real_oil_price, type = c("drift"), lags = 2)
summary(DF_real_oil_price_inter)

PP_real_oil_price = ur.pp(data_var$real_oil_price, type = "Z-tau", model = "constant", use.lag = 2)
summary(PP_real_oil_price)

## DF test global oil stocks

arima_oecd_oil_stock <- arima(data_var$oecd_oil_stock, order = c(13,0,0))
LjungBox(arima_oecd_oil_stock$residuals, lag=seq(1,25,1))

DF_oecd_oil_stock = ur.df(data_var$oecd_oil_stock, type = c("trend"), lags = 12)
summary(DF_oecd_oil_stock)

PP_oecd_oil_stock = ur.pp(data_var$oecd_oil_stock, type = "Z-tau", model = "trend", use.lag = 12)
summary(PP_oecd_oil_stock)

## DF test EA ind. prod.

arima_ind_prod_ea <- arima(data_var$ind_prod_ea, order = c(3,0,0))
LjungBox(arima_ind_prod_ea$residuals, lag=seq(1,25,1))

DF_ind_prod_ea = ur.df(data_var$ind_prod_ea, type = c("trend"), lags = 2)
summary(DF_ind_prod_ea)

DF_ind_prod_ea_inter = ur.df(data_var$ind_prod_ea, type = c("drift"), lags = 2)
summary(DF_ind_prod_ea_inter)

DF_ind_prod_ea_none = ur.df(data_var$ind_prod_ea, type = c("none"), lags = 2)
summary(DF_ind_prod_ea_none)

## DF test EA CPI

arima_cpi_ea <- arima(data_var$cpi_ea, order = c(15,0,0))
LjungBox(arima_cpi_ea$residuals, lag=seq(1,25,1))

DF_cpi_ea = ur.df(data_var$cpi_ea, type = c("trend"), lags = 14)
summary(DF_cpi_ea)

DF_cpi_ea_inter = ur.df(data_var$cpi_ea, type = c("drift"), lags = 14)
summary(DF_cpi_ea_inter)

DF_cpi_ea_none = ur.df(data_var$cpi_ea, type = c("none"), lags = 14)
summary(DF_cpi_ea_none)

PP_cpi_ea = ur.pp(data_var$cpi_ea, type = "Z-tau", model = "trend", use.lag = 14)
summary(PP_cpi_ea)

PP_cpi_ea_2 = PP.test(data_var$cpi_ea)
PP_cpi_ea_2$p.value

arima_cpi_ea_lv <- arima(data$ea19_cpi, order = c(20,0,0))
LjungBox(arima_cpi_ea_lv$residuals, lag=seq(1,25,1))

DF_cpi_ea_lv = ur.df(data$ea19_cpi, type = c("trend"), lags = 19)
summary(DF_cpi_ea_lv)

```


First differences transformation on unit root tests

```{r}
oil_prod_diff <- diff(data_var$oil_prod, lag = 1)
world_ind_pn_diff <- diff(data_var$world_ind_prod, lag = 1)
ind_prod_ea_diff <- diff(data_var$ind_prod_ea, lag = 1)
cpi_ea_diff <- diff(data_var$cpi_ea, lag =1)

data_var_diff <- data_var[-1,]

data_var_diff$oil_prod_diff <- oil_prod_diff
data_var_diff$world_ind_prod_diff <- world_ind_pn_diff
data_var_diff$ind_prod_ea_diff <- ind_prod_ea_diff
data_var_diff$cpi_ea_diff <- cpi_ea_diff

data_var_diff <- subset(data_var_diff, select = - c(oil_prod, world_ind_prod, ind_prod_ea, cpi_ea))  

order = c("oil_prod_diff","world_ind_prod_diff","real_oil_price","oecd_oil_stock","ind_prod_ea_diff", "cpi_ea_diff", "date")
data_var_diff <- data_var_diff[, order]
```

Graphs in log - diff 

```{r}

oil_prod_diff_g <- ggplot(data_var_diff, aes(x=date, y=oil_prod_diff)) +
  geom_line(color = "navy", size = 1) +
  theme_bw() +
  labs(x="Date", y="World oil prod. (MMBD)") +
  theme(text=element_text(size=12,  family="serif"))

world_ind_prod_diff_g <- ggplot(data_var_diff, aes(x=date, y=world_ind_prod_diff)) +
  geom_line(color = "navy", size = 1) +
  theme_bw() +
  labs(x="Date", y="World ind. prod. (index)") +
  theme(text=element_text(size=12,  family="serif"))

ind_prod_ea_diff_g <- ggplot(data_var_diff, aes(x=date, y=ind_prod_ea_diff)) +
  geom_line(color = "navy", size = 1) +
  theme_bw() +
  labs(x="Date", y="EA Ind. prod. (index)") +
  theme(text=element_text(size=12,  family="serif"))

cpi_ea_diff_g <- ggplot(data_var_diff, aes(x=date, y=cpi_ea_diff)) +
  geom_line(color = "navy", size = 1) +
  theme_bw() +
  labs(x="Date", y="EA CPI (index)") +
  theme(text=element_text(size=12,  family="serif"))

oil_market_diff_g <- ggarrange(oil_prod_diff_g, world_ind_prod_diff_g, ncol = 1, nrow = 2)
macro_diff_g <- ggarrange(ind_prod_ea_diff_g, cpi_ea_diff_g, ncol = 1, nrow = 2)

oil_market_diff_g
macro_diff_g
```



Estimation in log-diff and log-levels on unit root tests. Period Period 1986:1 - 2022:12

```{r}
y <- cbind(data_var_diff$oil_prod_diff, data_var_diff$world_ind_prod_diff, data_var_diff$real_oil_price, data_var_diff$oecd_oil_stock, data_var_diff$ind_prod_ea_diff, data_var_diff$cpi_ea_diff)

colnames(y) <- c("oil_prod", "world_ind_prod", "real_oil_price", "glob_oil_stock", "ind_prod_ea", "cpi_ea")

## period 1986:1 - 2022:12

y <- y[-1:-125,]

oil_supply_surprise <- data.frame(data$oil_supply_surprise)

oil_supply_surprise <- oil_supply_surprise[-1:-144,]

z <- cbind(oil_supply_surprise)
colnames(z) <- c("oil_supply_surprise")


## VAR using IV method
# Step 1

reduced_form_VAR <- VAR(y, p = 12, type = "const")

#Instantaneous response using instrument


shockcol <- externalinstrument(reduced_form_VAR, z, "real_oil_price")
shockcol


## IRFs

ma_representation <- Phi(reduced_form_VAR, 50)

irfs <- apply(ma_representation, 3, function(x) x %*% shockcol)

irfs <- as.data.frame(t(irfs))

colnames(irfs) <- names(shockcol)

irfs <- mutate(irfs, horizon = 0:50)

irfs <- gather(irfs, key = variable, value = response, -horizon)


irfs_oil_prod <- irfs[irfs$variable == 'oil_prod',]
lower_irfs_oil_prod = irfs_oil_prod$response - sd(irfs_oil_prod$response) #Lower 95% limit
upper_irfs_oil_prod = irfs_oil_prod$response + sd(irfs_oil_prod$response) #Upper 95% limit

irfs_world_ind_prod <- irfs[irfs$variable == 'world_ind_prod',]
lower_irfs_world_ind_prod = irfs_world_ind_prod$response - sd(irfs_world_ind_prod$response) #Lower 95% limit
upper_irfs_world_ind_prod = irfs_world_ind_prod$response + sd(irfs_world_ind_prod$response) #Upper 95% limit

irfs_real_oil_price <- irfs[irfs$variable == 'real_oil_price',]
lower_irfs_real_oil_price = irfs_real_oil_price$response - sd(irfs_real_oil_price$response) #Lower 95% limit
upper_irfs_real_oil_price = irfs_real_oil_price$response + sd(irfs_real_oil_price$response) #Upper 95% limit

irfs_glob_oil_stock <- irfs[irfs$variable == 'glob_oil_stock',]
lower_irfs_glob_oil_stock = irfs_glob_oil_stock$response - sd(irfs_glob_oil_stock$response) #Lower 95% limit
upper_irfs_glob_oil_stock = irfs_glob_oil_stock$response + sd(irfs_glob_oil_stock$response) #Upper 95% limit

irfs_ind_prod_ea <- irfs[irfs$variable == 'ind_prod_ea',]
lower_irfs_ind_prod_ea = irfs_ind_prod_ea$response - sd(irfs_ind_prod_ea$response) #Lower 95% limit
upper_irfs_ind_prod_ea = irfs_ind_prod_ea$response + sd(irfs_ind_prod_ea$response) #Upper 95% limit

irfs_cpi_ea <- irfs[irfs$variable == 'cpi_ea',]
lower_irfs_cpi_ea = irfs_cpi_ea$response - sd(irfs_cpi_ea$response) #Lower 95% limit
upper_irfs_cpi_ea = irfs_cpi_ea$response + sd(irfs_cpi_ea$response) #Upper 95% limit


irfs_oil_prod_g <- ggplot(irfs_oil_prod, aes(x = horizon, y = response)) +
  geom_line(size = 1, col = "black") +
  geom_line(aes(y = lower_irfs_oil_prod), color = "red", linetype = 5) +
  geom_line(aes(y = upper_irfs_oil_prod), color = "red", linetype = 5) +
  ggtitle("World oil production") +
  theme_bw() +
  labs(x="Horizon", y="Percent change (MMBD)") +
  theme(text=element_text(size=10,  family="serif"))


irfs_world_ind_prod_g <- ggplot(irfs_world_ind_prod, aes(x = horizon, y = response)) +
  geom_line(size = 1, col = "black") +
  geom_line(aes(y = lower_irfs_world_ind_prod), color = "red", linetype = 5) +
  geom_line(aes(y = upper_irfs_world_ind_prod), color = "red", linetype = 5) +
  ggtitle("World industrial production") +
  theme_bw() +
  labs(x="Horizon", y="Percent change (index)") +
  theme(text=element_text(size=10,  family="serif"))

irfs_real_oil_price_g <- ggplot(irfs_real_oil_price, aes(x = horizon, y = response)) +
  geom_line(size = 1, col = "black") +
  geom_line(aes(y = lower_irfs_real_oil_price), color = "red", linetype = 5) +
  geom_line(aes(y = upper_irfs_real_oil_price), color = "red", linetype = 5) +
  ggtitle("Real oil price") +
  theme_bw() +
  labs(x="Horizon", y="Percent (USD)") +
  theme(text=element_text(size=10,  family="serif"))

irfs_glob_oil_stock_g <- ggplot(irfs_glob_oil_stock, aes(x = horizon, y = response)) +
  geom_line(size = 1, col = "black") +
  geom_line(aes(y = lower_irfs_glob_oil_stock), color = "red", linetype = 5) +
  geom_line(aes(y = upper_irfs_glob_oil_stock), color = "red", linetype = 5) +
  ggtitle("World oil inventories") +
  theme_bw() +
  labs(x="Horizon", y="Percent (MMB)") +
  theme(text=element_text(size=10,  family="serif"))

irfs_ind_prod_ea_g <- ggplot(irfs_ind_prod_ea, aes(x = horizon, y = response)) +
  geom_line(size = 1, col = "black") +
  geom_line(aes(y = lower_irfs_ind_prod_ea), color = "red", linetype = 5) +
  geom_line(aes(y = upper_irfs_ind_prod_ea), color = "red", linetype = 5) +
  ggtitle("EA industrial production") +
  theme_bw() +
  labs(x="Horizon", y="Percent change (index)") +
  theme(text=element_text(size=10,  family="serif"))

irfs_cpi_ea_g <- ggplot(irfs_cpi_ea, aes(x = horizon, y = response)) +
  geom_line(size = 1, col = "black") +
  geom_line(aes(y = lower_irfs_cpi_ea), color = "red", linetype = 5) +
  geom_line(aes(y = upper_irfs_cpi_ea), color = "red", linetype = 5) +
  ggtitle("EA CPI") +
  theme_bw() +
  labs(x="Horizon", y="Percent change (index)") +
  theme(text=element_text(size=10,  family="serif"))


## Cum sum for IRFs in log diff

# oil prod

irfs_cumsum_oil_prod <- cumsum(irfs_oil_prod$response)
irfs_cumsum_oil_prod <- data.frame(irfs_cumsum_oil_prod)
irfs_cumsum_oil_prod$horizon <- irfs_oil_prod$horizon
colnames(irfs_cumsum_oil_prod) <- c("response", "horizon")

lower_irfs_cumsum_oil_prod = irfs_cumsum_oil_prod$response - sd(irfs_cumsum_oil_prod$response) #Lower 95% limit
upper_irfs_cumsum_oil_prod = irfs_cumsum_oil_prod$response + sd(irfs_cumsum_oil_prod$response) #Upper 95% limit

irfs_cum_oil_prod_g <- ggplot(irfs_cumsum_oil_prod, aes(x = horizon, y = response)) +
  geom_line(size = 1, col = "black") +
  geom_line(aes(y = lower_irfs_cumsum_oil_prod), color = "red", linetype = 5) +
  geom_line(aes(y = upper_irfs_cumsum_oil_prod), color = "red", linetype = 5) +
  ggtitle("World oil production") +
  theme_bw() +
  labs(x="Horizon", y="Cum. percent change (MMBD)") +
  theme(text=element_text(size=10,  family="serif"))

# World ind prod

irfs_cumsum_world_ind_prod <- cumsum(irfs_world_ind_prod$response)
irfs_cumsum_world_ind_prod <- data.frame(irfs_cumsum_world_ind_prod)
irfs_cumsum_world_ind_prod$horizon <- irfs_world_ind_prod$horizon
colnames(irfs_cumsum_world_ind_prod) <- c("response", "horizon")

lower_irfs_cumsum_world_ind_prod = irfs_cumsum_world_ind_prod$response - sd(irfs_cumsum_world_ind_prod$response) #Lower 95% limit
upper_irfs_cumsum_world_ind_prod = irfs_cumsum_world_ind_prod$response + sd(irfs_cumsum_world_ind_prod$response) #Upper 95% limit

irfs_cum_world_ind_prod_g <- ggplot(irfs_cumsum_world_ind_prod, aes(x = horizon, y = response)) +
  geom_line(size = 1, col = "black") +
  geom_line(aes(y = lower_irfs_cumsum_world_ind_prod), color = "red", linetype = 5) +
  geom_line(aes(y = upper_irfs_cumsum_world_ind_prod), color = "red", linetype = 5) +
  ggtitle("World industrial production") +
  theme_bw() +
  labs(x="Horizon", y="Cum. percent change (index)") +
  theme(text=element_text(size=10,  family="serif"))

# EA ind prod

irfs_cumsum_ind_prod_ea <- cumsum(irfs_ind_prod_ea$response)
irfs_cumsum_ind_prod_ea <- data.frame(irfs_cumsum_ind_prod_ea)
irfs_cumsum_ind_prod_ea$horizon <- irfs_ind_prod_ea$horizon
colnames(irfs_cumsum_ind_prod_ea) <- c("response", "horizon")

lower_irfs_cumsum_ind_prod_ea = irfs_cumsum_ind_prod_ea$response - sd(irfs_cumsum_ind_prod_ea$response) #Lower 95% limit
upper_irfs_cumsum_ind_prod_ea = irfs_cumsum_ind_prod_ea$response + sd(irfs_cumsum_ind_prod_ea$response) #Upper 95% limit

irfs_cum_ind_prod_ea_g <- ggplot(irfs_cumsum_ind_prod_ea, aes(x = horizon, y = response)) +
  geom_line(size = 1, col = "black") +
  geom_line(aes(y = lower_irfs_cumsum_ind_prod_ea), color = "red", linetype = 5) +
  geom_line(aes(y = upper_irfs_cumsum_ind_prod_ea), color = "red", linetype = 5) +
  ggtitle("EA industrial production") +
  theme_bw() +
  labs(x="Horizon", y="Cum. percent change (index)") +
  theme(text=element_text(size=10,  family="serif"))

# EA cpi

irfs_cumsum_cpi_ea <- cumsum(irfs_cpi_ea$response)
irfs_cumsum_cpi_ea <- data.frame(irfs_cumsum_cpi_ea)
irfs_cumsum_cpi_ea$horizon <- irfs_cpi_ea$horizon
colnames(irfs_cumsum_cpi_ea) <- c("response", "horizon")

lower_irfs_cumsum_cpi_ea = irfs_cumsum_cpi_ea$response - sd(irfs_cumsum_cpi_ea$response) #Lower 95% limit
upper_irfs_cumsum_cpi_ea = irfs_cumsum_cpi_ea$response + sd(irfs_cumsum_cpi_ea$response) #Upper 95% limit

irfs_cum_cpi_ea_g <- ggplot(irfs_cumsum_cpi_ea, aes(x = horizon, y = response)) +
  geom_line(size = 1, col = "black") +
  geom_line(aes(y = lower_irfs_cumsum_cpi_ea), color = "red", linetype = 5) +
  geom_line(aes(y = upper_irfs_cumsum_cpi_ea), color = "red", linetype = 5) +
  ggtitle("EA CPI") +
  theme_bw() +
  labs(x="Horizon", y="Cum. percent change (index)") +
  theme(text=element_text(size=10,  family="serif"))

```
IC for lag length 

```{r}

#Lag selection 

lagSelect = VARselect(y, lag.max = 24, type = "const")
stargazer(lagSelect$criteria, type = "latex")

```


Graphs for results chapter in document

```{r}

# Cum percent change (log - diff) variables: oil prod, world ind prod, EA ind prod, EA cpi

irfs_cum_oil_prod_g
irfs_cum_world_ind_prod_g
irfs_cum_ind_prod_ea_g
irfs_cum_cpi_ea_g

# Percent (log - level) variables: real oil price, global oil stocks

irfs_real_oil_price_g
irfs_glob_oil_stock_g


# Results

results_1_g <- ggarrange(irfs_cum_oil_prod_g, irfs_cum_world_ind_prod_g, ncol = 1, nrow = 2)
results_2_g <- ggarrange(irfs_real_oil_price_g, irfs_glob_oil_stock_g, ncol = 1, nrow = 2)
results_3_g <- ggarrange(irfs_cum_ind_prod_ea_g, irfs_cum_cpi_ea_g, ncol = 1, nrow = 2)

```

Robustness check

```{r} 

## Estimation using 24 lags 

## VAR using IV method
# Step 1

reduced_form_VAR_24 <- VAR(y, p = 24, type = "const")

#Instantaneous response using instrument


shockcol_24 <- externalinstrument(reduced_form_VAR_24, z, "real_oil_price")
shockcol_24


## IRFs

ma_representation_24 <- Phi(reduced_form_VAR_24, 50)

irfs_24 <- apply(ma_representation_24, 3, function(x) x %*% shockcol_24)

irfs_24 <- as.data.frame(t(irfs_24))

colnames(irfs_24) <- names(shockcol_24)

irfs_24 <- mutate(irfs_24, horizon = 0:50)

irfs_24 <- gather(irfs_24, key = variable, value = response, -horizon)


irfs_oil_prod_24 <- irfs_24[irfs_24$variable == 'oil_prod',]
lower_irfs_oil_prod_24 = irfs_oil_prod_24$response - sd(irfs_oil_prod_24$response) #Lower 95% limit
upper_irfs_oil_prod_24 = irfs_oil_prod_24$response + sd(irfs_oil_prod_24$response) #Upper 95% limit

irfs_world_ind_prod_24 <- irfs_24[irfs_24$variable == 'world_ind_prod',]
lower_irfs_world_ind_prod_24 = irfs_world_ind_prod_24$response - sd(irfs_world_ind_prod_24$response) #Lower 95% limit
upper_irfs_world_ind_prod_24 = irfs_world_ind_prod_24$response + sd(irfs_world_ind_prod_24$response) #Upper 95% limit

irfs_real_oil_price_24 <- irfs_24[irfs_24$variable == 'real_oil_price',]
lower_irfs_real_oil_price_24 = irfs_real_oil_price_24$response - sd(irfs_real_oil_price_24$response) #Lower 95% limit
upper_irfs_real_oil_price_24 = irfs_real_oil_price_24$response + sd(irfs_real_oil_price_24$response) #Upper 95% limit

irfs_glob_oil_stock_24 <- irfs_24[irfs_24$variable == 'glob_oil_stock',]
lower_irfs_glob_oil_stock_24 = irfs_glob_oil_stock_24$response - sd(irfs_glob_oil_stock_24$response) #Lower 95% limit
upper_irfs_glob_oil_stock_24 = irfs_glob_oil_stock_24$response + sd(irfs_glob_oil_stock_24$response) #Upper 95% limit

irfs_ind_prod_ea_24 <- irfs_24[irfs_24$variable == 'ind_prod_ea',]
lower_irfs_ind_prod_ea_24 = irfs_ind_prod_ea_24$response - sd(irfs_ind_prod_ea_24$response) #Lower 95% limit
upper_irfs_ind_prod_ea_24 = irfs_ind_prod_ea_24$response + sd(irfs_ind_prod_ea_24$response) #Upper 95% limit

irfs_cpi_ea_24 <- irfs_24[irfs_24$variable == 'cpi_ea',]
lower_irfs_cpi_ea_24 = irfs_cpi_ea_24$response - sd(irfs_cpi_ea_24$response) #Lower 95% limit
upper_irfs_cpi_ea_24 = irfs_cpi_ea_24$response + sd(irfs_cpi_ea_24$response) #Upper 95% limit

## Graphs

irfs_real_oil_price_g_24 <- ggplot(irfs_real_oil_price_24, aes(x = horizon, y = response)) +
  geom_line(size = 1, col = "black") +
  geom_line(aes(y = lower_irfs_real_oil_price_24), color = "red", linetype = 5) +
  geom_line(aes(y = upper_irfs_real_oil_price_24), color = "red", linetype = 5) +
  ggtitle("Real oil price") +
  theme_bw() +
  labs(x="Horizon", y="Percent (USD)") +
  theme(text=element_text(size=10,  family="serif"))

irfs_glob_oil_stock_g_24 <- ggplot(irfs_glob_oil_stock_24, aes(x = horizon, y = response)) +
  geom_line(size = 1, col = "black") +
  geom_line(aes(y = lower_irfs_glob_oil_stock_24), color = "red", linetype = 5) +
  geom_line(aes(y = upper_irfs_glob_oil_stock_24), color = "red", linetype = 5) +
  ggtitle("World oil inventories") +
  theme_bw() +
  labs(x="Horizon", y="Percent (MMB)") +
  theme(text=element_text(size=10,  family="serif"))

## Cum sum for IRFs in log diff

# oil prod

irfs_cumsum_oil_prod_24 <- cumsum(irfs_oil_prod_24$response)
irfs_cumsum_oil_prod_24 <- data.frame(irfs_cumsum_oil_prod_24)
irfs_cumsum_oil_prod_24$horizon <- irfs_oil_prod_24$horizon
colnames(irfs_cumsum_oil_prod_24) <- c("response", "horizon")

lower_irfs_cumsum_oil_prod_24 = irfs_cumsum_oil_prod_24$response - sd(irfs_cumsum_oil_prod_24$response) #Lower 95% limit
upper_irfs_cumsum_oil_prod_24 = irfs_cumsum_oil_prod_24$response + sd(irfs_cumsum_oil_prod_24$response) #Upper 95% limit

irfs_cum_oil_prod_g_24 <- ggplot(irfs_cumsum_oil_prod_24, aes(x = horizon, y = response)) +
  geom_line(size = 1, col = "black") +
  geom_line(aes(y = lower_irfs_cumsum_oil_prod_24), color = "red", linetype = 5) +
  geom_line(aes(y = upper_irfs_cumsum_oil_prod_24), color = "red", linetype = 5) +
  ggtitle("World oil production") +
  theme_bw() +
  labs(x="Horizon", y="Cum. percent change (MMBD)") +
  theme(text=element_text(size=10,  family="serif"))

# World ind prod

irfs_cumsum_world_ind_prod_24 <- cumsum(irfs_world_ind_prod_24$response)
irfs_cumsum_world_ind_prod_24 <- data.frame(irfs_cumsum_world_ind_prod_24)
irfs_cumsum_world_ind_prod_24$horizon <- irfs_world_ind_prod_24$horizon
colnames(irfs_cumsum_world_ind_prod_24) <- c("response", "horizon")

lower_irfs_cumsum_world_ind_prod_24 = irfs_cumsum_world_ind_prod_24$response - sd(irfs_cumsum_world_ind_prod_24$response) #Lower 95% limit
upper_irfs_cumsum_world_ind_prod_24 = irfs_cumsum_world_ind_prod_24$response + sd(irfs_cumsum_world_ind_prod_24$response) #Upper 95% limit

irfs_cum_world_ind_prod_g_24 <- ggplot(irfs_cumsum_world_ind_prod_24, aes(x = horizon, y = response)) +
  geom_line(size = 1, col = "black") +
  geom_line(aes(y = lower_irfs_cumsum_world_ind_prod_24), color = "red", linetype = 5) +
  geom_line(aes(y = upper_irfs_cumsum_world_ind_prod_24), color = "red", linetype = 5) +
  ggtitle("World industrial production") +
  theme_bw() +
  labs(x="Horizon", y="Cum. percent change (index)") +
  theme(text=element_text(size=10,  family="serif"))

# EA ind prod

irfs_cumsum_ind_prod_ea_24 <- cumsum(irfs_ind_prod_ea_24$response)
irfs_cumsum_ind_prod_ea_24 <- data.frame(irfs_cumsum_ind_prod_ea_24)
irfs_cumsum_ind_prod_ea_24$horizon <- irfs_ind_prod_ea_24$horizon
colnames(irfs_cumsum_ind_prod_ea_24) <- c("response", "horizon")

lower_irfs_cumsum_ind_prod_ea_24 = irfs_cumsum_ind_prod_ea_24$response - sd(irfs_cumsum_ind_prod_ea_24$response) #Lower 95% limit
upper_irfs_cumsum_ind_prod_ea_24 = irfs_cumsum_ind_prod_ea_24$response + sd(irfs_cumsum_ind_prod_ea_24$response) #Upper 95% limit

irfs_cum_ind_prod_ea_g_24 <- ggplot(irfs_cumsum_ind_prod_ea_24, aes(x = horizon, y = response)) +
  geom_line(size = 1, col = "black") +
  geom_line(aes(y = lower_irfs_cumsum_ind_prod_ea_24), color = "red", linetype = 5) +
  geom_line(aes(y = upper_irfs_cumsum_ind_prod_ea_24), color = "red", linetype = 5) +
  ggtitle("EA industrial production") +
  theme_bw() +
  labs(x="Horizon", y="Cum. percent change (index)") +
  theme(text=element_text(size=10,  family="serif"))

# EA cpi

irfs_cumsum_cpi_ea_24 <- cumsum(irfs_cpi_ea_24$response)
irfs_cumsum_cpi_ea_24 <- data.frame(irfs_cumsum_cpi_ea_24)
irfs_cumsum_cpi_ea_24$horizon <- irfs_cpi_ea_24$horizon
colnames(irfs_cumsum_cpi_ea_24) <- c("response", "horizon")

lower_irfs_cumsum_cpi_ea_24 = irfs_cumsum_cpi_ea_24$response - sd(irfs_cumsum_cpi_ea_24$response) #Lower 95% limit
upper_irfs_cumsum_cpi_ea_24 = irfs_cumsum_cpi_ea_24$response + sd(irfs_cumsum_cpi_ea_24$response) #Upper 95% limit

irfs_cum_cpi_ea_g_24 <- ggplot(irfs_cumsum_cpi_ea_24, aes(x = horizon, y = response)) +
  geom_line(size = 1, col = "black") +
  geom_line(aes(y = lower_irfs_cumsum_cpi_ea_24), color = "red", linetype = 5) +
  geom_line(aes(y = upper_irfs_cumsum_cpi_ea_24), color = "red", linetype = 5) +
  ggtitle("EA CPI") +
  theme_bw() +
  labs(x="Horizon", y="Cum. percent change (index)") +
  theme(text=element_text(size=10,  family="serif"))

# Graphs for document

results_1_g_24 <- ggarrange(irfs_cum_oil_prod_g_24, irfs_cum_world_ind_prod_g_24, ncol = 1, nrow = 2)
results_2_g_24 <- ggarrange(irfs_real_oil_price_g_24, irfs_glob_oil_stock_g_24, ncol = 1, nrow = 2)
results_3_g_24 <- ggarrange(irfs_cum_ind_prod_ea_g_24, irfs_cum_cpi_ea_g_24, ncol = 1, nrow = 2)


```


```{r} 

## Estimation using log levels 

y_l <- cbind(data_var$oil_prod, data_var$world_ind_prod, data_var$real_oil_price, data_var$oecd_oil_stock, data_var$ind_prod_ea, data_var$cpi_ea)

colnames(y_l) <- c("oil_prod", "world_ind_prod", "real_oil_price", "glob_oil_stock", "ind_prod_ea", "cpi_ea")

## period 1986:1 - 2022:12

y_l <- y_l[-1:-126,]

oil_supply_surprise_l <- data.frame(data$oil_supply_surprise)

oil_supply_surprise_l <- oil_supply_surprise_l[-1:-144,]

z_l <- cbind(oil_supply_surprise_l)
colnames(z_l) <- c("oil_supply_surprise")


## VAR using IV method
# Step 1

reduced_form_VAR_l <- VAR(y_l, p = 12, type = "const")

#Instantaneous response using instrument


shockcol_l <- externalinstrument(reduced_form_VAR_l, z_l, "real_oil_price")
shockcol_l


## IRFs

ma_representation_l <- Phi(reduced_form_VAR_l, 50)

irfs_l <- apply(ma_representation_l, 3, function(x) x %*% shockcol_l)

irfs_l <- as.data.frame(t(irfs_l))

colnames(irfs_l) <- names(shockcol_l)

irfs_l <- mutate(irfs_l, horizon = 0:50)

irfs_l <- gather(irfs_l, key = variable, value = response, -horizon)


irfs_oil_prod_l <- irfs_l[irfs_l$variable == 'oil_prod',]
lower_irfs_oil_prod_l = irfs_oil_prod_l$response - sd(irfs_oil_prod_l$response) #Lower 95% limit
upper_irfs_oil_prod_l = irfs_oil_prod_l$response + sd(irfs_oil_prod_l$response) #Upper 95% limit

irfs_world_ind_prod_l <- irfs_l[irfs_l$variable == 'world_ind_prod',]
lower_irfs_world_ind_prod_l = irfs_world_ind_prod_l$response - sd(irfs_world_ind_prod_l$response) #Lower 95% limit
upper_irfs_world_ind_prod_l = irfs_world_ind_prod_l$response + sd(irfs_world_ind_prod_l$response) #Upper 95% limit

irfs_real_oil_price_l <- irfs_l[irfs_l$variable == 'real_oil_price',]
lower_irfs_real_oil_price_l = irfs_real_oil_price_l$response - sd(irfs_real_oil_price_l$response) #Lower 95% limit
upper_irfs_real_oil_price_l = irfs_real_oil_price_l$response + sd(irfs_real_oil_price_l$response) #Upper 95% limit

irfs_glob_oil_stock_l <- irfs_l[irfs_l$variable == 'glob_oil_stock',]
lower_irfs_glob_oil_stock_l = irfs_glob_oil_stock_l$response - sd(irfs_glob_oil_stock_l$response) #Lower 95% limit
upper_irfs_glob_oil_stock_l = irfs_glob_oil_stock_l$response + sd(irfs_glob_oil_stock_l$response) #Upper 95% limit

irfs_ind_prod_ea_l <- irfs_l[irfs_l$variable == 'ind_prod_ea',]
lower_irfs_ind_prod_ea_l = irfs_ind_prod_ea_l$response - sd(irfs_ind_prod_ea_l$response) #Lower 95% limit
upper_irfs_ind_prod_ea_l = irfs_ind_prod_ea_l$response + sd(irfs_ind_prod_ea_l$response) #Upper 95% limit

irfs_cpi_ea_l <- irfs_l[irfs_l$variable == 'cpi_ea',]
lower_irfs_cpi_ea_l = irfs_cpi_ea_l$response - sd(irfs_cpi_ea_l$response) #Lower 95% limit
upper_irfs_cpi_ea_l = irfs_cpi_ea_l$response + sd(irfs_cpi_ea_l$response) #Upper 95% limit

# Graphs

irfs_oil_prod_g_l <- ggplot(irfs_oil_prod_l, aes(x = horizon, y = response)) +
  geom_line(size = 1, col = "black") +
  geom_line(aes(y = lower_irfs_oil_prod_l), color = "red", linetype = 5) +
  geom_line(aes(y = upper_irfs_oil_prod_l), color = "red", linetype = 5) +
  ggtitle("World oil production") +
  theme_bw() +
  labs(x="Horizon", y="Percent (MMBD)") +
  theme(text=element_text(size=10,  family="serif"))


irfs_world_ind_prod_g_l <- ggplot(irfs_world_ind_prod_l, aes(x = horizon, y = response)) +
  geom_line(size = 1, col = "black") +
  geom_line(aes(y = lower_irfs_world_ind_prod_l), color = "red", linetype = 5) +
  geom_line(aes(y = upper_irfs_world_ind_prod_l), color = "red", linetype = 5) +
  ggtitle("World industrial production") +
  theme_bw() +
  labs(x="Horizon", y="Percent (index)") +
  theme(text=element_text(size=10,  family="serif"))

irfs_real_oil_price_g_l <- ggplot(irfs_real_oil_price_l, aes(x = horizon, y = response)) +
  geom_line(size = 1, col = "black") +
  geom_line(aes(y = lower_irfs_real_oil_price_l), color = "red", linetype = 5) +
  geom_line(aes(y = upper_irfs_real_oil_price_l), color = "red", linetype = 5) +
  ggtitle("Real oil price") +
  theme_bw() +
  labs(x="Horizon", y="Percent (USD)") +
  theme(text=element_text(size=10,  family="serif"))

irfs_glob_oil_stock_g_l <- ggplot(irfs_glob_oil_stock_l, aes(x = horizon, y = response)) +
  geom_line(size = 1, col = "black") +
  geom_line(aes(y = lower_irfs_glob_oil_stock_l), color = "red", linetype = 5) +
  geom_line(aes(y = upper_irfs_glob_oil_stock_l), color = "red", linetype = 5) +
  ggtitle("World oil inventories") +
  theme_bw() +
  labs(x="Horizon", y="Percent (MMB)") +
  theme(text=element_text(size=10,  family="serif"))

irfs_ind_prod_ea_g_l <- ggplot(irfs_ind_prod_ea_l, aes(x = horizon, y = response)) +
  geom_line(size = 1, col = "black") +
  geom_line(aes(y = lower_irfs_ind_prod_ea_l), color = "red", linetype = 5) +
  geom_line(aes(y = upper_irfs_ind_prod_ea_l), color = "red", linetype = 5) +
  ggtitle("EA industrial production") +
  theme_bw() +
  labs(x="Horizon", y="Percent (index)") +
  theme(text=element_text(size=10,  family="serif"))

irfs_cpi_ea_g_l <- ggplot(irfs_cpi_ea_l, aes(x = horizon, y = response)) +
  geom_line(size = 1, col = "black") +
  geom_line(aes(y = lower_irfs_cpi_ea_l), color = "red", linetype = 5) +
  geom_line(aes(y = upper_irfs_cpi_ea_l), color = "red", linetype = 5) +
  ggtitle("EA CPI") +
  theme_bw() +
  labs(x="Horizon", y="Percent (index)") +
  theme(text=element_text(size=10,  family="serif"))

# Graphs for document

results_1_g_l <- ggarrange(irfs_oil_prod_g_l, irfs_world_ind_prod_g_l, ncol = 1, nrow = 2)
results_2_g_l <- ggarrange(irfs_real_oil_price_g_l, irfs_glob_oil_stock_g_l, ncol = 1, nrow = 2)
results_3_g_l <- ggarrange(irfs_ind_prod_ea_g_l, irfs_cpi_ea_g_l, ncol = 1, nrow = 2)


```